\large{\bf Looking forwards and backwards: dynamics and genealogies of locally regulated populations }

### 2023-12-15 review-responses.tex

Thanks ETCETERA

\subsection*{Major comments}

\begin{point}{\revref}
The first paragraph seems sloppy. For example, this expression [for $\mu_\theta(x)$] is non-negative for
what range of values for $(x, \eta)$? Why ``anyway true for sufficiently large $\theta$''? Perhaps this
assumption should be labeled and called upon whenever is it being referred to.
\end{point}

\reply{
    We've revised this to remove the $\max$ in Equation~\eqref{eqn:mu_defn},
    and make the assumption that this is nonnegative more consistently:
    see \revref{} and the new paragraph following Assumptions~\ref{def:model_setup}.
}

\begin{point}{}
    Definition \ref{def:lineage_generator}, Proposition \ref{thm:lineages}:
    I think these may need to be rewritten/rephrased. Up
to Proposition \ref{thm:lineages}, it is not clear what ``that individual'' is in the limiting model. Therefore,
it is not clear what $L_s$ is. While the definition of the process $L^N$ maybe clear, what is the
precise definition of $L_s$ ? I think these need to be defined through the finite-$N$ models. It may
be good to explain why the process $L$ is Markovian, or at least point out that it is Markovian,
before talking about semigroup. Probably Section \ref{sec:lineage_motion} will need to be modified accordingly.
\end{point}

\reply{
    This is a bit tricky, because really to justify this being a sensible notion
    we'd like to refer to the lookdown, which has not yet been introduced.
    To make things more clear we've added a bit to the earlier introductory paragraph
    \llname{whatis_lineage},
    reworded Definition~\ref{def:lineage_generator} somewhat,
    and added some more explanation after the Definition and before the Proposition,
    which also points out that it is Markovian \revref.
}

@@

\begin{point}{\revref}
The two conditions in equation \eqref{conditions for one step} are crucial for the one-step convergence.
It would be great to mention exactly where each of them are used in Section \ref{subsec:one step convergence proof}.
A brief explanation about these conditions would also be nice. For example, it seems the first (high
migration/mixing rate) condition in \eqref{conditions for one step},
but not the second (high density) condition, is used
to guarantees that $\psi_t^{\epsilon,x}$ is close to a heat kernel?
\end{point}

\reply{
    We've added a bit more here \revref,
    and also referred the reader back to the (at this point long-ago)
    discussion earlier \llname{one_step_heuristics}.
}

@@

\subsection*{Minor comments}

\begin{point}{\revref}
\eqref{eqn:mu_defn}: The authors mentioned ``define the death rate'' and ``where this equation defines F'',
which is confusing. What is being defined here? Please clarify. It seems to me that one has
to first choose $F$, then define $\mu_\theta(x)$ as in \eqref{eqn:mu_defn}.
\end{point}

\reply{
    That's right; we've rearranged this so it's hopefully more clear \revref.
}

@@

\begin{point}{\revref}
% P7, paragraph after Remark 2.3:
``and once there establishes...'' It maybe better to split this
long sentence into two complete sentences.
\end{point}

\reply{
    We've rearranged this to be shorter  \revref.
}

\begin{point}{\revref}
% P7, paragraph after Remark 2.3:
$\meanq(x)$ and $\covq(x)$ did not appear before. Perhaps mention they
are given/chosen, or simply delete them here, and mention them in Definition \ref{def:dispersal_generator}.
\end{point}

\reply{
    That's right, this bit was supposed to introduce those functions
    (and we want to keep them here to emphasize the scaling with $1/\theta$).
    We've rewritten this to make this more clear \revref.
}

@@

\begin{point}{\revref}
Remark \ref{ancestral lineages: first guess}
Why use $m$ in the superscript of $q_\theta^m(x, \eta, dy)$?
\end{point}

\reply{
    We've added a clarifying point \revref.
}

\begin{point}{}
Assumptions \ref{def:model_setup}: Should it be ``Assumption 2.8'' instead? Also, $\alpha$ in condition 1 was
used in $\alpha = lim_{N \to \infty} \theta(N)/N$ on Page 8, Figure 1, etc.
\end{point}

\reply{
    Good question -- it could go either way,
    but since there's more than one assumption, we'd like to keep it plural?
    And, good point about re-use of $\alpha$; we've changed it to 
    ``$\beta$-H\"older continuous for some $\beta \in (0, 1]$''.
}

@@

\begin{point}{\revref}
% P9 last line:
    Should the ``$C$'' on the right hand side of the equal sign [in the equation for $C_f$] be deleted ?
\end{point}

\reply{
    The notation there was supposed to convey what exactly about $f$ the constant $C_f$
    depends on, i.e., it could be defined via a function $C : \IR_{\ge 0} \to \IR_{\ge 0}$
    whose argument is the stuff in the parentheses.
    We could write this out explicitly, but think it's fairly standard to write things this way
    -- if on re-reading it is still confusing, we're happy to change it, though?
}

@@

\begin{point}{\revref}
% P10 paragraph below Lemma 2.9:
    Why ``no longer suffices'', and why this is true? . Please
explain. I don’t seem to find this explained in any remark before this.
\end{point}

\reply{
    We intended the \emph{next} sentence to explain why this is true;
    to make this more clear
    we changed that sentence from
    \emph{``Perhaps the simplest example in which we can see this is the case''}
    to
    \emph{``Perhaps the simplest example in which we can see
    that non-local reproduction can lead to rapid growth even when $r\gamma=\mu$''}
    \revref.
}

\begin{point}{}
    Figure \ref{fig:super_vs_det_2d}: What is $N$, the time and other parameters (e.g. $\rho_r$ , $\rho_\gamma$) in the simulations? Same
for other figures. Please at least specify the range of values.
\end{point}

\reply{
    We've added some more detail to the caption.
}

@@

\begin{point}{\revref}
% P12 paragraph after Corollary 2.11:
A more precise reference, or explanation for why
uniqueness holds for this special case would be helpful, as the references you ave are all
quite extensive.
\end{point}

\reply{
    Perhaps Evans/Perkins 1994 ``Measure-valued branching diffusions'' for Dawson's Girsanov thm?
    (reproduced as Thm IV.1.6 in Perkins, "Dawson-Watanabe Superprocesses...").
}

@@

\begin{point}{}
Proposition \ref{nonlocalPME to PME}:
    It seems the constant $C$ is not needed. Perhaps just say the supremum
of the first displayed expression over $\epsilon$ finite?
\end{point}

\reply{
    Good point! Done.
}

\begin{point}{}
Remark \ref{remark_on_nonlocal_to_local}: Can you explain a bit more why it will be necessary to retain averaging of r?
\end{point}

\reply{
    Good idea; done \revref.
}

\begin{point}{}
P16: Second paragraph:
The notation $L_t$ seems to first appear here and was not yet defined.
\end{point}

\reply{
    Ah, true - now we introduce it (albeit still informally) earlier \llname{Linf_mention}.
}

@@

\begin{point}{}
This maybe out of the scope of this paper, but it is related. As the authors pointed out, two
lineages probably will not coalesce in the limiting model. However, they will coalesce in
the finite-$N$ model, though it may take a very long time. I wonder if the authors have a
feeling about whether the description of the lineage dynamics in the limit still be a good
approximation for that of the finite model for such a long time.
\end{point}

\reply{
    Just say we don't know, but:
    see Pennington and Etheridge on bistable waves
    where fluctuations in the wave happen on long time scales
}

\begin{point}{}
The subtitle of section \ref{sec:nonunique_lineage} does not match the actual content in that section. Please
update the subtitle. Also, can you give a reference for the claim in the last sentence in this
section \revref?
\end{point}

\reply{
    We've retitled this section, and added a citation \revref.
}

@@

\begin{point}{\revref}
specify ``this process'' in the line below the first displayed equation in
Section 5.1.
\end{point}

\reply{
    Done \revref.
}

\begin{point}{\revref}
Paragraph starting with ``Levels never cross below 0'': Should it be instead ``Levels never
go above $N$''? If so, what particles are regarded as dead? Perhaps those exit the interval
$[0, N]$? Note that equation \eqref{eqn:dot_u} suggests that levels can now go above $N$.
\end{point}

\reply{
    This was correct as written;
    we've rephased it to be more clear \revref.
}

@@

\begin{point}{\revref}
% P32 third line in Section 5.3:
    Pointwise in time?
\end{point}

\reply{
    We mean for each $\omega$; now clarified \revref.
}

\begin{point}{\revref}
Proposition \ref{prop:limiting_construction}: In ``For any solution to the equation above'', what is a ``solution''? Is it
a solution for all time? Also, be specific to ``the equations above'' by labeling and referring
to the precise equations including the unlabeled displayed equation below \eqref{eqn:limiting_construction}.
\end{point}

\reply{
    Good suggestion; we've clarified, see \revref and \llname{XU_clarification}.
}

\begin{point}{}
The word ``Proof'' is missing in ``Proof of Proposition 5.6''. This word is also
missing in other places in Sections 6-8.
\end{point}

\reply{
    Whoops! That was due to a \LaTeX format change; now fixed.
}

@@

\begin{point}{}
``Proof of Proposition \ref{prop:limiting_construction}'': Do you also need the unlabeled displayed equation below
\eqref{eqn:limiting_construction}? The solution is for all time.
\end{point}

\reply{
    We think this was an unedited duplicate of point 1.24 above?
}

@@

\begin{point}{\revref}
Be specific about ``these results'' in ``Proof of the remainder of these results...'' Also, at the
very beginning of Remark \ref{remark_on_branching}, Be specific about what ``The process'' is.
\end{point}

\reply{
    Happily, there were no remaining results that we hadn't already said where the proofs were,
    so this line is deleted. We also said that ``the process'' is $\lp$ \revref.
}

@@

\begin{point}{\revref}
``Condition 1'' here is that for Lemma 2.9, not that for Assumptions 2.8. Please specify.
\end{point}

\reply{
    Good point - these were not called ``conditions'' in Lemma~\ref{lem:conditions on r},
    so we now refer to them as ``assumptions'' \revref.
}

@@

\begin{point}{\revref}
% P40 first displayed equation:
Mention $M_t$ is a martingale for all such test functions.
\end{point}

\reply{
    Done \revref.
}

\begin{point}{\revref}
%    P48:
    The notation $\mathcal{O}(\epsilon)$ seems to first appear here. Please explain its meaning. Do they
depend on x?
\end{point}

\reply{
    We've made the bounds here (and below) more explicit.
    \revref
}

@@

\begin{point}{\revref}
% P54: near the bottom:
    Can you be more specific what in paper \citet{barlow/jacka/yor:1986} is being used here?
\end{point}

\reply{
    The fact is that we can bound the expectation of the quadratic variation by a constant times the expectation of the angle bracket process;
    rather than explain again here we refer to the previous place we used the same argument
    \revref.
}

\begin{point}{\revref}
% P60:
The notation $\mathcal{O}(\delta)$ seems to first appear here. Please explain its meaning. Do they
depend on $x$?
\end{point}

\reply{
    We've clarified this (but also note that the expression does not depend on $x$) \revref.
}

@@

\begin{point}{\revref}
% P65: Remark 8.1:
measure-values processes $\to$ measure-valued processes
\end{point}

\reply{
    Done. \revref
}

\begin{point}{}
Section \ref{sec:lineages_proof} may need to be modified also based on the second point in ``Major
comments'' above.
\end{point}

\reply{
    Thanks! We didn't think that any modifications were necessary, though.
}

### 2023-12-15 forwards-and-backwards.tex
addressed most of the comments

\comment{TODO: double-check $\langle \cdot \rangle$ versus $[\cdot]$}
\comment{TODO: check factor of 2}

@@

As one takes a journey, long or short, the landscape changes:
forests thicken or thin or change their composition;
even in flat plains
springtime grasslands host intergrading mosaics of different types of flowers.
The aim of this paper is to introduce and study
a broad class of mechanistic spatial models
that might describe how spatially heterogeneous populations live, die, and reproduce.
Questions that we (start to) address include:
How does population density change across space and time?
How might we learn about the underlying dynamics from genealogical or genetic data?
And, how does genetic ancestry spread across geography
when looking back through time in these populations?

Reproduction of individuals naturally leads to spatial branching process models,
including branching random walk, 
branching Brownian motion, and the Dawson-Watanabe superprocesses.  
However, as a result of the branching assumption (once born, individuals behave 
independently of one another), 
a population evolving according to any of these models will
either die out or grow without bound and, in so doing, can
develop clumps of arbitrarily large density and extent. 
Our starting point here is an individual-based model of a single species 
in continuous space in which
birth, death, and establishment may all depend on local population density
as well as on spatial location,
allowing for stable populations through density-dependent feedback.
The model generalizes those introduced to the ecology literature by
\citet{bolker/pacala:1997} and \citet{law/murrell/dieckmann:2003},
and subsequently studied, e.g., by \citet{etheridge:2004} and \citet{fournier2004microscopic}.

@@

The history of a natural population is often only accessible indirectly, 
through patterns of genetic diversity that have been laid down; from 
genetic data, one can try to infer the genealogical trees that relate 
individuals in a sample from the population, and these have been shaped 
by its history (see e.g., \citep{neigel1993application,kelleher2019inferring}).
It is therefore of interest to establish information about the distribution of genealogical trees 
under our population model, which we do with a lookdown construction. 
Lookdown constructions were first introduced in~\cite{donnelly/kurtz:1996} to provide 
a mechanism for retaining information about genealogical relationships between individuals
sampled from a population evolving according to the Moran model when passing to the 
infinite population limit.
Since then, they have been extended to a wide range of models. Of particular
relevance to our work here are the papers of~\cite{kurtz/rodrigues:2011} 
and~\cite{etheridge/kurtz:2019}, in which
lookdown constructions are provided for a wide variety of population models, including
spatially structured branching processes.

@@

We shall write $\mu_\theta(x, X/N)$ for the per-capita death rate of a mature individual
at $x$ in the population.
In order for the 
population density to change over timescales of order $\theta$, we should like the net 
per capita reproductive rate 
to scale as $1/\theta$. 
In classical models, in which $r$, $\gamma$, and $\mu$ are constant, 
this quantity is simply $r\gamma -\mu$. 
Here, because production of juveniles and their establishment are
mediated by population density measured relative to different points, 
the net reproductive rate will take a more complicated form. 
In particular, the total rate of production of mature offspring by an individual at $x$ will be
\begin{equation} \label{eqn:total_repro}
	\gamma\big(x,\smooth{\gamma}X(x)/N\big)\int r\big(y, \smooth{r}X(y)/N\big)q_{\theta}(x,dy).
\end{equation}
Nonetheless, it will be convenient to define the death rate $\mu_\theta$ in terms of 
its deviation from $r\gamma$. 
To this end, we define
the death rate of an individual at $x$ to be \revpoint{1}{4}
\begin{align} \label{eqn:mu_defn}
    \mu_\theta(x, X/N)
    =
    \max\left\{0, 
        r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N)
    \right\} ,
\end{align}
where this equation also introduces $F(x, m) : \IR^d \times \IR_{\ge 0} \to \IR$,
and $\rho_F$ is again a smooth kernel. 
The function $F$ is nearly
the net per capita reproductive rate, scaled by $\theta$,
and would be equal to it in a nonspatial model;
but, as can be seen from \eqref{eqn:total_repro},
differs because an offspring's establishment probability is measured at their new location
rather than that of their parent.
For the most part, we work with $F$ instead of $\mu_\theta$.

@@

In summary, at each time $t$, $\eta^N_t$ is purely atomic, consisting of atoms of mass $1/N$
(which are the individuals).
At instantaneous rate $\theta \gamma(x, \eta^N_t) N \eta^N_t(dx)$ an offspring of mass $1/N$ is 
produced at location $x$; it disperses to 
a location $y$ offset from $x$
by an independent Gaussian random variable with mean $\meanq(x) / \theta$ \revpoint{1}{6}
and covariance matrix $\covq(x) / \theta$,
and once there establishes instantaneously \revpoint{1}{5}
with probability $r(y, \eta^N_t)$, or else dies. 
At instantaneous rate $\theta \mu_\theta(x, \eta^N_t) N \eta^N_t(dx)$ an individual at location 
$x$ dies.
Note that the process 
$\left(\eta^N_t\right)_{t\ge 0}$, which
records numbers and locations of adult individuals, is just a scaled spatial birth and 
death process. If, for example, we insist that $\gamma(x,m)$ is bounded, then 
existence (and in particular non-explosion) is guaranteed by comparison with
a pure birth process. We do not dwell on this, as we shall require more stringent conditions
if we are to pass to the limit as $\theta$ and $N$ tend to infinity.

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    As above, we define the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although often we omit the dependence of $\meanq$ and $\covq$ on $x$).
    Furthermore, we define for $f \in C_b^\infty(\IR^d)$,
    \begin{align}
    \DG f(x)
        =
        \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    and denote the adjoint of $\DG$ by
    \begin{align*}
    \DG^* f(x)
        &=
        \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
	    % \label{drift in adjoint}
        % &=
        % \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
	    % -\nabla f(x)\cdot \meanq(x)-f(x)\nabla\cdot\meanq(x).
        &=
        \sum_{ij} C_{ij}(x) \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \left(
            \sum_j \partial_{x_j} C_{ij}(x) - \meanq_i(x)
        \right) \partial_{x_i} f(x)
        \\ & \qquad {}
        + \left(
            \sum_{ij} \partial_{x_i} \partial_{x_j} C_{ij}(x)
            -
            \sum_i \partial_{x_i} \meanq_i(x)
        \right) f(x) .
    \end{align*}
\end{definition}

@@

\begin{remark}
\label{ancestral lineages: first guess}
An equivalent way to describe the model would be to say that
when the state of the population is $\eta$,
an individual at $x$ gives birth at rate
$$
    \theta \gamma(x, \eta)
    \int r(y, \eta) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$q_\theta^\mathfrak{m}$ (with superscript $\mathfrak{m}$ because it is post-\textbf{m}ortality),
defined by: \revpoint{1}{7}
$$
    q_\theta^\mathfrak{m}(x,\eta,dy)
    :=
    \frac{
        r(y, \eta) q_\theta(x, dy)
    }{
        \int r(z, \eta) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
is biased towards regions of higher establishment probability.
For comparison with future results,
it is interesting to write down the limiting generator:
\begin{align}
	\label{weighted dispersal}
    \lim_{\theta \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, \eta, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \eta) \right](x)
        - 
        f(x) \DG\left[ r(\cdot, \eta) \right](x)
    }{
        r(x, \eta)
    } .
\end{align}
In the simplest case of unbiased isotropic dispersal
(i.e.,~$\meanq = 0$ and $\covq = I$), $\DG = \Delta$,
and so~(\ref{weighted dispersal}) is equal to
\begin{align*}
    \Delta f(x) + 2 \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
One might guess that the spatial motion described by following  
the ancestral lineage of an individual 
back through time
would be described (in the limit) by the adjoint of this generator.
However, we will see in Section~\ref{sec:ancestral lineages}
that this is not in fact the case.
\end{remark}

@@

{\bf Dispersal generator:}
We assume that
\begin{enumerate}
    \item $\meanq(x)$ and $\covq(x)$ are
        $\alpha$-H\"older continuous for some $\alpha \in (0, 1]$
        and uniformly bounded
        in each component, and
    % \item $\meanq(x)$ and $\covq(x)$ are $\alpha$-H\"older continuous in each component, for some $\alpha \in (0,1]$;
    \item the operator $\DG$ is uniformly strictly elliptic,
        i.e., $\inf_x \inf_{y:\|y\|=1} \sum_{ij} y_i C(x)_{ij} y_j > 0$.
\end{enumerate}

{\bf Reproduction parameters:}
We assume that 
\begin{enumerate}
            \addtocounter{enumi}{2}
\item The function $F(x,m)$ satisfies 
    \begin{enumerate}
    \item $F(x,m)$ is locally Lipschitz in $m$;
    \item $F(x,m)$ is uniformly bounded above (but not necessarily below);
    \item for each fixed $m$, $\sup_{x\in\IR^d}\sup_{k\leq m}|F(x,k)|<\infty$;
    \end{enumerate}
\item The functions $r(x,m)$, $\gamma(x,m)$ have bounded first and second 
    derivatives in both arguments;
    \label{r_gamma_derivs_condition}
\item $\gamma(x,m)$ is uniformly bounded;
    \label{gamma_bounded_condition}
\item For each $f \in C_b^2(\IR^d)$,
    there is a $C_f$ such that
    $$
        |\gamma(x, \eta) \theta \int 
        (r(y,\eta) f(y) - r(x, \eta) f(x))
        q_\theta(x, dy)| \le C_f (1 + |f(x)|)
    $$
    for all $x \in \IR^d$ and $\eta \in \measures$.
    Furthermore, $C_f$ only depends on the norm of the first two derivatives of $f$,
    i.e., 
    \[
        C_f = C(\sup_x \sup_{\|z\|=1} \max( \sum_i z_i \partial_{x_i} f(x), \sum_{ij} z_i z_j \partial_{x_i x_j} f(x))) .
    \] \label{gamma_B_condition} \revpoint{1}{9}
\end{enumerate}
To keep expressions manageable, we shall also assume that  \revpoint{1}{1}
\[
\mu_\theta(x) = r(x,\eta) \gamma(x,\eta) - \frac{1}{\theta} F(x,\eta),
\]
that is, this expression is non-negative so that there is no need to take the maximum
	with zero in~(\ref{eqn:mu_defn}). (This is anyway true for 
	sufficiently large $\theta$.)
\end{assumptions}

@@

Now that we have established what we can say about how 
population density changes with time,
we turn to results on ancestral lineages,
i.e., how genealogical ancestry can be traced back across the landscape.
Informally,
a \emph{lineage} $(L_t^N)_{t \ge 0}$,
begun at a spatial location $L_0^N = x$
where there is a focal individual in the present day
can be obtained by, for each time $t$,
setting $L_t^N$ to be the spatial location of the individual alive at time $t$
before the present from
whom the focal individual is descended.
Since in our model individuals have only one parent, this is unambiguous.
Although we did not explicitly retain such information,
it is clear that for finite $N$, since individuals are born one at a time, 
one could construct the lineage $(L_t^N)_{t=0}^T$
given the history of the population $(\eta^N_t)_{t = 0}^T$,
for each starting location to which $\eta^N_T$ assigns positive mass.
It is less clear, however, how to formally retain such information when we 
pass to the scaling limit.

>>>>

\comment{TODO: insert statement that one might be worried about $\eta_t$ being well-defined as a process, in e.g., Def'n 5.2 and following?}

@@

As one takes a journey, long or short, the landscape changes:
forests thicken or thin or change their composition;
even in flat plains,
springtime prairies host intergrading mosaics of different types of flowers.
The aim of this paper is to introduce and study
a broad class of mechanistic spatial models
that might describe how spatially heterogeneous populations live, die, and reproduce.
Questions that we (start to) address include:
How does population density change across space and time?
How might we learn about the underlying dynamics from genealogical or genetic data?
And, how does genetic ancestry spread across geography
when looking back through time in these populations?

Reproduction of individuals naturally leads to spatial branching process models,
including branching random walk, 
branching Brownian motion, and the Dawson-Watanabe superprocesses.  
However, as a result of the branching assumption (once born, individuals behave 
independently of one another), 
a population evolving according to any of these models will
either die out or grow without bound and, in so doing, can
develop clumps of arbitrarily large density and extent. 
Our starting point here is an individual-based model of a single species 
in continuous space in which
birth, death, and establishment may all depend on local population density
as well as on spatial location,
allowing for stable populations through density-dependent feedback.
The model generalizes those introduced to the ecology literature by
\citet{bolker/pacala:1997} and \citet{law/murrell/dieckmann:2003},
and our work follows various others in the mathematical literature
(e.g., \citet{etheridge:2004,fournier2004microscopic}).

@@

The history of a natural population is often only accessible indirectly, 
through patterns of genetic diversity:
from genetic data, one can try to infer the genealogical trees that describe how
individuals in a sample from the population are related, and these have been shaped 
by its history (see e.g., \citep{neigel1993application,kelleher2019inferring}).
It is therefore of interest to establish information about the distribution of genealogical trees 
under our population model, which we do with a lookdown construction. 
Lookdown constructions were first introduced in~\cite{donnelly/kurtz:1996} to provide 
a mechanism for retaining information about genealogical relationships between individuals
sampled from a population evolving according to the Moran model when passing to the 
infinite population limit.
Since then, they have been extended to a wide range of models. Of particular
relevance to our work here are the papers of~\cite{kurtz/rodrigues:2011} 
and~\cite{etheridge/kurtz:2019}, in which
lookdown constructions are provided for a wide variety of population models, including
spatially structured branching processes.

@@

We shall write $\mu_\theta(x, X/N)$ for the per-capita death rate of a mature individual
at $x$ in the population.
In order for the 
population density to change over timescales of order $\theta$, we should like the net 
per capita reproductive rate 
to scale as $1/\theta$. 
In classical models, in which $r$, $\gamma$, and $\mu$ are constant, 
this quantity is simply $r\gamma -\mu$. 
Here, because production of juveniles and their establishment are
mediated by population density measured relative to different points, 
the net reproductive rate will take a more complicated form. 
In particular, the total rate of production of mature offspring by an individual at $x$ will be
\begin{equation} \label{eqn:total_repro}
	\gamma\big(x,\smooth{\gamma}X(x)/N\big)\int r\big(y, \smooth{r}X(y)/N\big)q_{\theta}(x,dy).
\end{equation}
Nonetheless, it will be convenient to define the death rate $\mu_\theta$ in terms of 
its deviation from $r\gamma$. 
To this end, we define
the death rate of an individual at $x$,  \revpoint{1}{4}
using a function $F(x, m) : \IR^d \times \IR_{\ge 0} \to \IR$,
as
\begin{align} \label{eqn:mu_defn}
    \mu_\theta(x, X/N)
    =
        r(x, \smooth{r} X(x) / N) \gamma(x, \smooth{\gamma} X(x) / N)
        - \frac{1}{\theta} F(x, \smooth{F} X(x) / N)
    ,
\end{align}
where $\rho_F$ is again a smooth kernel. 
(We will also assume that parameters are chosen so that this is always nonnegative,
a point we return to later.)
The function $F$ is nearly
the net per capita reproductive rate, scaled by $\theta$,
and would be equal to it in a nonspatial model;
but, as can be seen from \eqref{eqn:total_repro},
differs because an offspring's establishment probability is measured at their new location
rather than that of their parent.
For the most part, we work with $F$ instead of $\mu_\theta$.

@@

In summary, at each time $t$, $\eta^N_t$ is purely atomic, consisting of atoms of mass $1/N$
(which are the individuals).
At instantaneous rate $\theta \gamma(x, \eta^N_t) N \eta^N_t(dx)$ an offspring of mass $1/N$ is 
produced at location $x$, which
immediately disperses to a location $y$ offset from $x$
by an independent Gaussian random variable,
and once there establishes instantaneously \revpoint{1}{5}
with probability $r(y, \eta^N_t)$, or else dies. 
The distribution of the dispersal displacement (i.e., $y-x$) may depend on $x$,
and is specified by functions defining the mean $\meanq(x) / \theta$ \revpoint{1}{6}
and covariance matrix $\covq(x) / \theta$.
At instantaneous rate $\theta \mu_\theta(x, \eta^N_t) N \eta^N_t(dx)$ an individual at location 
$x$ dies.
Note that the process 
$\left(\eta^N_t\right)_{t\ge 0}$, which
records numbers and locations of adult individuals, is just a scaled spatial birth and 
death process. If, for example, we insist that $\gamma(x,m)$ is bounded, then 
existence (and in particular non-explosion) is guaranteed by comparison with
a pure birth process. We do not dwell on this, as we shall require more stringent conditions
if we are to pass to the limit as $\theta$ and $N$ tend to infinity.

@@

\begin{definition}[Dispersal generator]
    \label{def:dispersal_generator}
    As above, we define the dispersal kernel,
    $q_\theta(x, dy)$,
    to be the density of a multivariate Gaussian
    with mean $\meanq(x)/\theta$ and covariance matrix $\covq(x)/\theta$
    (although often we omit the dependence of $\meanq$ and $\covq$ on $x$).
    Furthermore, we define for $f \in C_b^\infty(\IR^d)$,
    \begin{align}
    \DG f(x)
        =
        \frac{1}{2} \sum_{ij} \covq(x)_{ij} \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \meanq(x)_i \partial_{x_i} f(x)
    \end{align}
    and denote the adjoint of $\DG$ by
    \begin{align*}
    \DG^* f(x)
        &=
        \frac{1}{2} \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
        - \sum_i \partial_{x_i} (f(x) \meanq(x)_i) 
        \\
	    % \label{drift in adjoint}
        % &=
        % \frac{1}{2} \sum_{ij} \partial_{x_i} \partial_{x_j} (\covq(x)_{ij} f(x))
	    % -\nabla f(x)\cdot \meanq(x)-f(x)\nabla\cdot\meanq(x).
        &=
        \frac{1}{2} \sum_{ij} C_{ij}(x) \partial_{x_i} \partial_{x_j} f(x)
        + \sum_i \left(
            \frac{1}{2} \sum_j \partial_{x_j} C_{ij}(x) - \meanq_i(x)
        \right) \partial_{x_i} f(x)
        \\ & \qquad {}
        + \left(
            \frac{1}{2} \sum_{ij} \partial_{x_i} \partial_{x_j} C_{ij}(x)
            -
            \sum_i \partial_{x_i} \meanq_i(x)
        \right) f(x) .
    \end{align*}
\end{definition}

@@

\begin{remark}
\label{ancestral lineages: first guess}
An equivalent way to describe the model would be to say that
when the state of the population is $\eta$,
an individual at $x$ gives birth at rate
$$
    \theta \gamma(x, \eta)
    \int r(y, \eta) q(x, dy) ,
$$
and that offspring disperse according to the kernel
$q_\theta^\mathfrak{m}$ (with superscript $\mathfrak{m}$ because it is post-\textbf{m}ortality),
defined by: \revpoint{1}{7}
$$
    q_\theta^\mathfrak{m}(x,\eta,dy)
    :=
    \frac{
        r(y, \eta) q_\theta(x, dy)
    }{
        \int r(z, \eta) q_\theta(x, dz)
    } .
$$
Clearly, the random walk driven by this dispersal kernel
is biased towards regions of higher establishment probability.
For comparison with future results,
it is interesting to write down the limiting generator:
\begin{align}
	\label{weighted dispersal}
    \lim_{\theta \to \infty}
    \theta \int (f(y) - f(x)) q_\theta^\mathfrak{m}(x, \eta, dy)
    &=
    \frac{
        \DG\left[ f(\cdot) r(\cdot, \eta) \right](x)
        - 
        f(x) \DG\left[ r(\cdot, \eta) \right](x)
    }{
        r(x, \eta)
    } .
\end{align}
In the simplest case of unbiased isotropic dispersal
(i.e.,~$\meanq = 0$ and $\covq = I$), $\DG = \Delta/2$,
and so~(\ref{weighted dispersal}) is equal to
\begin{align*}
    \frac{1}{2} \Delta f(x) + \grad f(x) \cdot \grad \log r(\cdot, \smooth{r} \eta(\cdot))(x) .
\end{align*}
One might guess that the spatial motion described by following  
the ancestral lineage of an individual 
back through time
would be described (in the limit) by the adjoint of this generator.
However, we will see in Section~\ref{sec:ancestral lineages}
that this is not in fact the case.
\end{remark}

@@

{\bf Dispersal generator:}
We assume that
\begin{enumerate}
    \item $\meanq(x)$ and $\covq(x)$ are
        $\beta$-H\"older continuous for some $\beta \in (0, 1]$
        and uniformly bounded
        in each component, and
    % \item $\meanq(x)$ and $\covq(x)$ are $\alpha$-H\"older continuous in each component, for some $\alpha \in (0,1]$;
    \item the operator $\DG$ is uniformly strictly elliptic,
        i.e., $\inf_x \inf_{y:\|y\|=1} \sum_{ij} y_i C(x)_{ij} y_j > 0$.
\end{enumerate}

{\bf Reproduction parameters:}
We assume that 
\begin{enumerate}
            \addtocounter{enumi}{2}
\item The function $F(x,m)$ satisfies 
    \begin{enumerate}
    \item $F(x,m)$ is locally Lipschitz in $m$;
    \item $F(x,m)$ is uniformly bounded above (but not necessarily below);
    \item for each fixed $m$, $\sup_{x\in\IR^d}\sup_{k\leq m}|F(x,k)|<\infty$;
    \end{enumerate}
\item The functions $r(x,m)$, $\gamma(x,m)$ have bounded first and second 
    derivatives in both arguments;
    \label{r_gamma_derivs_condition}
\item $\gamma(x,m)$ is uniformly bounded;
    \label{gamma_bounded_condition}
\item For each $f \in C_b^2(\IR^d)$,
    there is a $C_f$ such that
    $$
        |\gamma(x, \eta) \theta \int 
        (r(y,\eta) f(y) - r(x, \eta) f(x))
        q_\theta(x, dy)| \le C_f (1 + |f(x)|)
    $$
    for all $x \in \IR^d$ and $\eta \in \measures$.
    Furthermore, $C_f$ only depends on the norm of the first two derivatives of $f$,
    i.e., 
    \[
        C_f = C(\sup_x \sup_{\|z\|=1} \max( \sum_i z_i \partial_{x_i} f(x), \sum_{ij} z_i z_j \partial_{x_i x_j} f(x))) .
    \] \label{gamma_B_condition} \revpoint{1}{9}
\item To keep expressions manageable, we shall also assume that  \revpoint{1}{1}
    the death rate (as defined in~\eqref{eqn:mu_defn}) is nonnegative, i.e., that
    \[
    \mu_\theta(x) = r(x,\eta) \gamma(x,\eta) - \frac{1}{\theta} F(x,\eta) \ge 0 .
    \]
\end{enumerate}
\end{assumptions}

Since $F$ is bounded above,
the final assumption that $\mu_\theta \ge 0$ --
or, equivalently, that $F(x,\eta) \le \theta r(x,\eta) \gamma(x,\eta)$ --
will always be true, for instance, for large enough $\theta$ 
as long as $r$ and $\gamma$ are bounded away from zero.

@@

Now that we have established what we can say about how 
population density changes with time,
we turn to results on ancestral lineages,
i.e., how genealogical ancestry can be traced back across the landscape.
Informally,
a \emph{lineage} $(L_t^N)_{t \ge 0}$,
begun at a spatial location $L_0^N = x$
where there is a focal individual in the present day,
can be obtained for each time $t$ by
setting $L_t^N$ to be the spatial location of the individual alive at time $t$
before the present from
whom the focal individual is descended.
Since in our model individuals have only one parent, this is unambiguous.
Although we did not explicitly retain such information,
it is clear that for finite $N$, since individuals are born one at a time, 
one could construct the lineage $(L_t^N)_{t=0}^T$
given the history of the population $(\eta^N_t)_{t = 0}^T$,
for each starting location to which $\eta^N_T$ assigns positive mass.
It is less clear, however, how to formally retain such information when we 
pass to the scaling limit.
